---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

### setup the timezone
```{r}
Sys.setlocale("LC_ALL", "English")
```

###Load the packages needed:
```{r results="hide"}
library(lubridate)
library(dplyr)
```

## Loading and preprocessing the data

###Load activity data
```{r load data,echo=TRUE}
act<-read.csv("activity.csv")
```



## What is mean total number of steps taken per day?
### 1. Calculate the total number of steps taken per day and remove NAs
```{r}
act2<-act%>%group_by(date)%>%summarise(TotalStep=sum(steps))
rmna.act2<-act2[complete.cases(act2),]
head(rmna.act2)     ##showing head of the mean table (NAs removed)
```

### 2.Make a histogram of the total number of steps taken each day
```{r}
hist(rmna.act2$TotalStep,main="Total number of steps taken per day",xlab="Sum of steps",breaks=20,col="Dark Red")
```

### 3. Calculate and report the mean and median of the total number of steps taken per day
```{r}
summary(act2$TotalStep)
```



## What is the average daily activity pattern?
### 1. Make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
act3=mutate(act,interval=sprintf("%04d",interval)) ##make 5 to 0005
act3$interval=paste(substr(act3$interval,1,2),substr(act3$interval,3,4), sep=":") ## make HHMM to HH:MM
meanact3<-act3%>%group_by(interval)%>%summarise(meanstep=mean(steps,na.rm=TRUE))  ## calculate the mean
meanact3$interval=strptime(meanact3$interval,"%H:%M")     ##convert character to POSIXLt
plot(meanact3,type="l",xlab="Interval",ylab="Mean number of steps", main="Average Daily Activity Pattern",col="BLUE")
```

### 2. The maximum number of steps of all 5-minute interval
```{r}
meanact3$interval[meanact3$meanstep==max(meanact3$meanstep)]
```



## Imputing missing values
### 1. Calculate and report the total number of missing values in the dataset
```{r}
sum(is.na(act$steps))
```

### 2. Filling in all of the missing values in the dataset
```{r}
fillna<-act3
fillna$meanstep=rep(meanact3$meanstep,61)     ##Add meanstep to table
my.na<-is.na(fillna$steps)
fillna$steps[my.na]<-fillna$meanstep[my.na]  ## fill NA with meanstep per day
fillna$meanstep<-NULL  ## remove the column
head(fillna)     ##showing the head of table (NA filled with the mean for that 5-minute interval)

```

### 3. Make a histogram of the total number of steps taken each day
```{r}
act4<-fillna%>%group_by(date)%>%summarise(TotalStep=sum(steps))
hist(act4$TotalStep,main="Total number of steps taken per day",xlab="Sum of steps",breaks=20,col="Dark Red")
```

### 4. Calculate and report the mean and median total number of steps taken per day
```{r}
summary(act4$TotalStep)
```



## Are there differences in activity patterns between weekdays and weekends?
### 1. Create a new factor variable with ¡§weekday¡¨ and ¡§weekend¡¨
```{r}
act5<-fillna[1:3]
act5$wday<-weekdays(as.Date(act5$date))
weekday<-c("Monday","Tuesday","Wednesday","Thursday","Friday")
weekend<-c("Saturday","Sunday")
act5$wday[act5$wday%in%weekday]<-"weekday"   ## Replace Mon. to Fri. with "weekday"
act5$wday[act5$wday%in%weekend]<-"weekend"   ## Replace Sat. and Sun. with "weekend"
head(act5)   ## Showing the head of table
```

### 2. Make a panel plot
```{r}
wky<-act5[act5$wday=="weekday",]
wkd<-act5[act5$wday=="weekend",]
mwky<-wky%>%group_by(interval)%>%summarise(meanstep=mean(steps))
mwkd<-wkd%>%group_by(interval)%>%summarise(meanstep=mean(steps))
mwky$interval=strptime(mwky$interval,"%H:%M")     ##convert character to POSIXLt
mwkd$interval=strptime(mwkd$interval,"%H:%M")     ##convert character to POSIXLt
par(mfrow=c(2,1))
plot(mwky,type="l",xlab="Interval",ylab="Mean number of steps", main="Averaged Weekday Activity Pattern",col="BLUE")
plot(mwkd,type="l",xlab="Interval",ylab="Mean number of steps", main="Averaged Weekend Activity Pattern",col="Red")
```